\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2018-08-07 paper/ms.tex
Working on the changes

\Abstract{In probabilistic (Bayesian) inferences, we typically want to
  compute properties of the posterior distribution, describing
  knowledge of unknown quantities in the context of a particular
  dataset and the assumed prior information. The marginal likelihood,
  also known as the ``evidence'', is a key quantity in Bayesian model
  selection. The diffusive nested sampling algorithm, a variant of
  nested sampling, is a powerful tool for generating posterior samples
  and estimating marginal likelihoods. It is effective at solving
  complex problems including many where the posterior distribution is
  multimodal or has strong dependencies between variables.
  \pkg{DNest4} is an open source (MIT licensed), multi-threaded
  implementation of this algorithm in \proglang{C++11}, along with
  associated utilities including: i) `\code{RJObject}', a class template
  for finite mixture models; (ii) a \proglang{Python} package allowing
  basic use without \proglang{C++} coding; and iii) experimental
  support for models implemented in \proglang{Julia}.  In this paper
  we demonstrate \pkg{DNest4} usage through examples including simple
  Bayesian data analysis, finite mixture models, and approximate
  Bayesian computation.  }

@@

In the nested sampling sphere,
the simple \proglang{C} implementation of given by
\citet{skilling2006nested} is useful for understanding the classic nested sampling
algorithm. Similarly to \pkg{DNest4}, it is the user's responsibility
to implement the MCMC moves used. The DNS algorithm produces more
accurate estimates of $Z$ and ought to outperform classic nested sampling on multimodal
problems.

@@

\section[Specifying models in Julia]{Specifying models in \proglang{Julia}}
Experimental support for model specification in the
\proglang{Julia} language \citep{julia}
is provided. Directory
\code{code/Templates/JuliaModel} contains an example.
Set an environment variable \code{JULIA\_LIB\_PATH} to the directory on your
system that contains the libraries \code{libjulia.so} and
\code{sys.so}. Set another environment variable \code{JULIA\_INCLUDE\_PATH}
to the directory where the \proglang{Julia} header files are located
(do not include the final \code{julia} directory in this path).

The log-likelihood is specified in \code{julia\_model.jl}, as a function
called \code{log\_likelihood} that takes a \code{Vector::Float64} as an argument
and returns a \code{Float64}. You should assume that the priors for the
parameters are independent Uniform(0, 1) distributions (sensible default
proposals are used). To use a different prior, transform the parameters inside
your \code{log\_likelihood} function before using them.
You may add other functions to \code{julia\_model.jl}, for example to load
a dataset to refer to in your \code{log\_likelihood}. The example in
the repository is the ``SpikeSlab'' likelihood from
\citet{brewer2011diffusive}.

You can compile the model by invoking \code{make}
from the \code{JuliaModel} directory.
Only a single thread is supported when running \proglang{Julia}
models.

>>>>

\Abstract{In probabilistic (Bayesian) inferences, we typically want to
  compute properties of the posterior distribution, describing
  knowledge of unknown quantities in the context of a particular
  dataset and the assumed prior information. The marginal likelihood,
  also known as the ``evidence'', is a key quantity in Bayesian model
  selection. The diffusive nested sampling algorithm, a variant of
  nested sampling, is a powerful tool for generating posterior samples
  and estimating marginal likelihoods. It is effective at solving
  complex problems including many where the posterior distribution is
  multimodal or has strong dependencies between variables.
  \pkg{DNest4} is an open source (MIT licensed), multi-threaded
  implementation of this algorithm in \proglang{C++11}, along with
  associated utilities including: i) `\code{RJObject}', a class template
  for finite mixture models; and (ii) a \proglang{Python} package allowing
  basic use without \proglang{C++} coding.  In this paper
  we demonstrate \pkg{DNest4} usage through examples including simple
  Bayesian data analysis, finite mixture models, and approximate
  Bayesian computation.  }

@@

In the nested sampling sphere,
the simple \proglang{C} implementation given by
\citet{skilling2006nested} is useful for understanding the classic nested sampling
algorithm. Similarly to \pkg{DNest4}, it is the user's responsibility
to implement the MCMC moves used. The DNS algorithm produces more
accurate estimates of $Z$ and ought to outperform classic nested sampling on multimodal
problems.

@@

%\section[Specifying models in Julia]{Specifying models in \proglang{Julia}}
%Experimental support for model specification in the
%\proglang{Julia} language \citep{julia}
%is provided. Directory
%\code{code/Templates/JuliaModel} contains an example.
%Set an environment variable \code{JULIA\_LIB\_PATH} to the directory on your
%system that contains the libraries \code{libjulia.so} and
%\code{sys.so}. Set another environment variable \code{JULIA\_INCLUDE\_PATH}
%to the directory where the \proglang{Julia} header files are located
%(do not include the final \code{julia} directory in this path).

%The log-likelihood is specified in \code{julia\_model.jl}, as a function
%called \code{log\_likelihood} that takes a \code{Vector::Float64} as an argument
%and returns a \code{Float64}. You should assume that the priors for the
%parameters are independent Uniform(0, 1) distributions (sensible default
%proposals are used). To use a different prior, transform the parameters inside
%your \code{log\_likelihood} function before using them.
%You may add other functions to \code{julia\_model.jl}, for example to load
%a dataset to refer to in your \code{log\_likelihood}. The example in
%the repository is the ``SpikeSlab'' likelihood from
%\citet{brewer2011diffusive}.

%You can compile the model by invoking \code{make}
%from the \code{JuliaModel} directory.
%Only a single thread is supported when running \proglang{Julia}
%models.
